---
title: "African-enculturated participants"
Me: Ségolène M. R. Guérin & Emmanuel Coulon
Date: "2026-02-12"
output:
  html_document:
    toc: yes
  pdf_document:
    toc: yes
  html_notebook:
    code_folding: hide
    toc: yes
  word_document:
    toc: yes
editor_options:
  markdown:
    wrap: sentence
---
This is the notebook associated to statistical tests for the African-enculturated participants (Stage 2 #1)

# Preamble
```{r preamble}
# ------ CLEANING R SESSION ####
rm(list=ls()) # clean environment window
graphics.off() # clean plot window

# ------ PACKAGE LOADING ####
library(writexl) # to export .xlxs file
library(readxl) # to read .xlxs file
library(here) # to load data in the project data folder
library(dplyr) # to use infix operator %>%
library(naniar) # to use replace_with_na_all function
library(pracma) # to have findpeaks functions
library(tidyr) # to tidy data
library(rstatix) # to detecte outliers
library(ggplot2) # to plot data
library(readxl) # to load excel file
library(ggpubr) # to build qq plot
library(lsr) # to compute Cohen's D for pairwise comparisons
library(matrixStats) # to compute SD over rows
library(readr) # to load tsv files
library(MetBrewer) # to have more palettes
library(ggrepel) # to have label on line chart
library(svglite) # to export .svg files
library(moments) # to compute skewness and kurtosis
library(sur) # to compute standardised skewness
library(GGally) # for parallel plots
library(lmerTest) # to build stat model
library(performance) # to run model checks
library(lmom)
library(DescTools) # to winsorize
library(emmeans)   # for post-hoc comparisons with the GLMs
library(TOSTER) # for TOST
library(bestNormalize)

# ------ FUNCTION AND EXTERN FILES CALL ####
source(here("r_functions/motor_parallel_plot.R"))
source(here("r_functions/plots.R"))
source(here("r_functions/calc_age.R"))
source(here("r_functions/guess_delim.R"))

# ------ PARAMETERS ####
# ------ Define analysis type
# 1 = zscore as described in PRR 
# 2 = normalized zscore                                                       (not used in final manuscript)
# 3 = normalized zscore corrected by the stim zscore
# 4 = normalized 2 metRel pour le four beat                                   (not used in final manuscript)
# 5 = 3beat (1.25, 2.5, 3.75) vs 4beat (1.66, 3.33)
# 6 = normalized 3beat (1.25, 2.5, 3.75) vs 4beat (1.66, 3.33)                (not used in final manuscript)
# 7 = normalized 3beat (1.25, 2.5, 3.75) vs 4beat (1.66, 3.33) - stim
# 8 = (3beat (1.25, 2.5, 3.75) vs 4beat (1.66, 3.33) - stim) * normalization  (not used in final manuscript)
analysis_type = 1


if (analysis_type == 1){
  dv = "zscore_freq_rel"
} else if (analysis_type == 2){
  dv = "norm_zscore_freq_rel"
} else if (analysis_type == 3){
  dv = "norm_zscore_freq_rel"
} else if (analysis_type == 4){
  dv = "zscore_freq_rel"
} else if (analysis_type == 5){
  dv = "zscore_freq_rel"
} else if (analysis_type == 6){
  dv = "norm_zscore_freq_rel"  
} else if (analysis_type == 7){
  dv = "norm_zscore_freq_rel"
} else if (analysis_type == 8){
  dv = "norm_zscore_freq_rel"  
}

# ------ Load participants file
participant_pool <- 
  read_excel(here("participants_final_pool.xlsx"))
```


# H1 – Neural data

## Load data

```{r}
# ------ H1 ####
# ------ DATA LOADING
# ------ Indicate data path
path_z_scores = 
  gsub("1_code/","", here("2_output/data/4_final/eeg/z_score_fft/")) 

# ------ List all files
list_files <- matrix(list.files(path = path_z_scores, pattern = "^grp"),
                    byrow = TRUE) 

# ------ Load data
data_file <- data.frame()

for (participant in 1:nrow(list_files)) {
   
  participant_file <-  read_csv(paste(path_z_scores, list_files[participant],
                                      sep = "/"),
                                show_col_types = FALSE)
   
   data_file <- rbind(data_file, participant_file)
}

# ------ Keep only relevant participants
data_file <-
  data_file[which(data_file$participant %in%
                    c(participant_pool$african_duple,
                      participant_pool$african_triple)), ]


# ------ Keep only relevant data
if (analysis_type == 1 || analysis_type == 3) { 
  # Only keep 2 meter-related frequencies in the four-beat meter
  data_file <- data_file %>%
    filter(    
      (condition == 2 & metrical_interp == 2) |
        (condition == 3 & metrical_interp == 3))
  
}  else if (analysis_type == 5 || analysis_type == 7) { 
  # 3beat (1.25 2.5 3.75 Hz) and 4beat (1.66 3.33 Hz)
  data_file <- data_file %>%
    filter(    
      (condition == 2 & metrical_interp == 23) |
        (condition == 3 & metrical_interp == 32))
}


# Subtract stimulus zscore value is needed
if (analysis_type == 3){
  for (participant in 1:nrow(data_file)){
    if (data_file$metrical_interp[participant] == 2){
      data_file$norm_zscore_freq_rel[participant] <- 
        data_file$norm_zscore_freq_rel[participant] - (-0.192572263128259)
    } else if (data_file$metrical_interp[participant] == 3){
      data_file$norm_zscore_freq_rel[participant] <- 
        data_file$norm_zscore_freq_rel[participant] - (-0.317879804131487)}}}

if (analysis_type == 7){
  for (participant in 1:nrow(data_file)){
    if (data_file$metrical_interp[participant] == 23){
      data_file$norm_zscore_freq_rel[participant] <- 
        data_file$norm_zscore_freq_rel[participant] - (-0.243056176623996)
    } else if (data_file$metrical_interp[participant] == 32){
      data_file$norm_zscore_freq_rel[participant] <- 
        data_file$norm_zscore_freq_rel[participant] - (-0.181114652805544)}}}
  
# ------ Prepare data
# Keep only relevant data
data_file <- data_file[ , c(1:4, 6:7)]

# Change data format
data_file[1:4] <- lapply(data_file[1:4], as.factor)
```


## Stats

### Descriptives

```{r}
# ------ STATISTICS
# ------ Descriptive
stat_descr <-
  data_file %>%
  group_by(condition, session) %>%
  get_summary_stats(dv,
                    type = "mean_sd")
print(stat_descr)

```

### Assumptions checks

Normality of residuals and homoscedasticity assumptions are met.
One outlier is detected.

```{r}
# ------ Assumptions check
model <- lmer(as.formula(paste(dv, "~ condition * session + (1 | participant)")), 
              data = data_file)

check_normality(model)
check_heteroscedasticity(model)

# ------ Identify outliers
check_outliers(model, method = "zscore")

```

#### Outliers correction

The outlier was corrected using Winsorisation (see [Leys et al., 2019](https://biblio.ugent.be/publication/8619726)).

```{r}
# ------ Deal with outliers
backup = data_file

if (analysis_type == 1 || analysis_type == 3 || analysis_type == 7) {

  if (analysis_type == 1) { 
    probs_min = 0
    probs_max = 0.98 
  } else if (analysis_type == 3 || analysis_type == 7) {    
    probs_min = 0
    probs_max = 0.99    
  }

  data_file[[dv]] <-
    Winsorize(data_file[[dv]],
              val = quantile(data_file[[dv]], probs = c(probs_min, probs_max)))
    
  # Print modified rows
  changed_rows <- data_file[[dv]] != backup[[dv]]
  
  changed_rows_number <- which(changed_rows)
  
  cbind(line = changed_rows_number, data_file[changed_rows, ])
  
  model <- lmer(as.formula(paste(dv, "~ condition * session + (1 | participant)")),
                data = data_file)
  check_normality(model)
  check_outliers(model, method = "zscore")
}


# save data_file for paper figure (with outlier correction)
path_outlier_z_scores =  gsub("1_code/","",
                              here("2_output/data/4_final/eeg/z_score_fft/outliers_correction/")) 
write.csv(data_file, 
          paste0(path_outlier_z_scores,"data_file_eeg_grp1_analysis_type",analysis_type,"_outliers_corrected.csv"))

```


### Inferential

No significant effect. 

```{r}
# ------ Inferential
# Omnibus test
data_file %>%
  anova_test(dv = dv, wid = participant,
             between = condition,
             within = session,
             effect.size = "pes") %>%
  get_anova_table()
   
# GLM
anova(model, type = 3) 
effectsize::eta_squared(model)

```


## Plot

```{r}
# ------ PLOT
# ------ Indicate export path
output_dir <- 
  gsub("1_code/", "",
       here("2_output/plots/eeg/"))

# stim zscores
  if (analysis_type == 1) {
    z_stim = c(-0.365, -0.369)
  } else if (analysis_type == 2) {
    z_stim = c(-0.193, -0.318)
  } else if (analysis_type == 3) {
    z_stim = c(0, 0)
  } else if (analysis_type == 4) {
    z_stim = c(-0.365, -0.344)
  } else if (analysis_type == 5) {
    z_stim = c(-0.352, -0.344)
  } else if (analysis_type == 6) {
    z_stim = c(-0.243, -0.181)
  } else if (analysis_type == 7) {
    z_stim = c(0, 0)
  } else if (analysis_type == 8) {
    z_stim = c(0, 0)    
  }


# ------ Build plot
ggplot(data_file,
       aes(y = .data[[dv]],
           x = condition,
           col = session)) +
  geom_point(aes(y = .data[[dv]]),
             alpha = 0.5,
             size = 2,
             position = position_jitterdodge(jitter.width = 0.15,
                                             jitter.height = 0,
                                             dodge.width = 0.3)) +
  geom_boxplot(width = 0.3,
               alpha = 0,
               lwd = 1) +
  geom_flat_violin(aes(fill = session),
                   position = position_nudge(x = .2, y = 0),
                   alpha = .5) +
  
  labs(y = "z scores",
       x = "Movement Condition",
       color = "Session",
       fill = "Session",
       subtitle = "EEG data") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray") +
  annotate("segment", x = 0.6, xend = 1.4, y = z_stim[1], yend = z_stim[1], color = "darkgray", alpha = 0.5) +
  annotate("segment", x = 1.6, xend = 2.4, y = z_stim[2], yend = z_stim[2], color = "darkgray", alpha = 0.5) +
  scale_color_manual(values = met.brewer("Java", 2)) +
  scale_fill_manual(values = met.brewer("Java", 2)) +
  scale_x_discrete(labels = c("Three beat", "Four beat")) +
  scale_y_continuous(name = "z scores", breaks = seq(from = -1, to = 1.5, by = 0.5), limits = c(-1.2,1.7)) +
  apa7

# Save
name_file <- c("african_eeg_zscores.png")
ggsave(paste(output_dir, name_file, sep = '/'), plot = last_plot(),
       dpi = 300, width = 35, height = 25, units = "cm", scale = 0.5)
```

# --------------------------------------
# H2 – Behaviour data

## Load data

```{r}
# ------ H2 ####
# ------ DATA LOADING
# ------ Indicate data path
path_z_scores = 
  gsub("1_code/","", here("2_output/data/4_final/clap/z_score/")) 

# ------ List all files
list_files <- matrix(list.files(path = path_z_scores, pattern = "^grp"),
                    byrow = TRUE) 

# ------ Load data
data_file <- data.frame()

for (participant in 1:nrow(list_files)) {
   
  participant_file <-  read_csv(paste(path_z_scores, list_files[participant],
                                      sep = "/"),
                                show_col_types = FALSE)
   
   data_file <- rbind(data_file, participant_file)
}

# Keep only relevant participants
data_file <-
  data_file[which(data_file$participant %in%
                    c(participant_pool$african_duple,
                      participant_pool$african_triple)), ]

# ------ Keep only relevant data
if (analysis_type == 1 || analysis_type == 3) { 
  data_file <- data_file %>%
    filter(    
      (condition == 2 & metrical_interp == 2) |
        (condition == 3 & metrical_interp == 3))
  
}  else if (analysis_type == 5 || analysis_type == 7) { 
  # 3beat (1.25 2.5 3.75 Hz) and 4beat (1.66 3.33 Hz)
  data_file <- data_file %>%
    filter(    
      (condition == 2 & metrical_interp == 23) |
        (condition == 3 & metrical_interp == 32))
}

# Subtract stimulus zscore value is needed
if (analysis_type == 3){
  for (participant in 1:nrow(data_file)){
    if (data_file$metrical_interp[participant] == 2){
      data_file$norm_zscore_freq_rel[participant] <- 
        data_file$norm_zscore_freq_rel[participant] - (-0.192572263128259)
    } else if (data_file$metrical_interp[participant] == 3){
      data_file$norm_zscore_freq_rel[participant] <- 
        data_file$norm_zscore_freq_rel[participant] - (-0.317879804131487)}}}


if (analysis_type == 7){
  for (participant in 1:nrow(data_file)){
    if (data_file$metrical_interp[participant] == 23){
      data_file$norm_zscore_freq_rel[participant] <- 
        data_file$norm_zscore_freq_rel[participant] - (-0.243056176623996)
    } else if (data_file$metrical_interp[participant] == 32){
      data_file$norm_zscore_freq_rel[participant] <- 
        data_file$norm_zscore_freq_rel[participant] - (-0.181114652805544)}}}

# ------ Prepare data
# Keep only relevant data
data_file <- data_file[ , c(1:4, 6:7)]

# Change data format
data_file[1:4] <- lapply(data_file[1:4], as.factor)

```


## Stats

### Descriptives

```{r}
# ------ STATISTICS
# ------ Descriptive
stat_descr <-
  data_file %>%
  group_by(condition, session) %>%
  get_summary_stats(dv, type = "mean_sd")
print(stat_descr)

```


### Assumptions checks

Normality of residuals and homoscedasticity assumptions are met.
No outliers are detected.

```{r}
# ------ Assumptions check
model <- lmer(as.formula(paste(dv, "~ condition * session + (1 | participant)")),
              data = data_file)

check_normality(model)
check_heteroscedasticity(model)

# ------ Identify outliers
check_outliers(model, method = "zscore")

# save data_file for paper figure
write.csv(data_file, paste0("/Users/emmanuelcoulon/Documents/MATLAB/PROJECTS/meterlearning_motor/2_output/plots/paper/data_file_clap_grp1_analysis_type",analysis_type,"_outliers_corrected.csv"))

```

### Inferential

Significant effect of session.

```{r}
# ------ Inferential
# Omnibus test
data_file %>%
  anova_test(dv = dv, wid = participant,
             between = condition,
             within = session,
             covariate = ,
             effect.size = "pes") %>%
  get_anova_table()

# GLM
anova(model, type = 3) 
effectsize::eta_squared(model)

```
#### H2a

zclapping will be larger after when compared to before movement.

```{r}

# Data summary 
data_file %>%
  group_by(session) %>%
  get_summary_stats(dv, type = "mean_sd")

# Post-hoc
data_file %>%
  t_test(as.formula(paste(dv, "~ session")), paired = TRUE,
                  p.adjust.method = "bonferroni")

# Effect size
data_file %>%
  cohens_d(as.formula(paste(dv, "~ session")),
           paired = TRUE)


# ----- GLM - Estimated Marginal Means
contrast_model <- emmeans(model, pairwise ~ session, adjust = "bonferroni")
print(contrast_model)
eff_size(contrast_model, sigma = sigma(model), edf = 38)

```

## Plot

```{r}
# ------ PLOT
# ------ Indicate export path
output_dir <- 
  gsub("1_code/", "",
       here("2_output/plots/clap/"))

# stim zscores
  if (analysis_type == 1) {
    z_stim = c(-0.365, -0.369)
  } else if (analysis_type == 2) {
    z_stim = c(-0.193, -0.318)
  } else if (analysis_type == 3) {
    z_stim = c(0, 0)
  } else if (analysis_type == 4) {
    z_stim = c(-0.365, -0.344)
  } else if (analysis_type == 5) {
    z_stim = c(-0.352, -0.344)
  } else if (analysis_type == 6) {
    z_stim = c(-0.243, -0.181)
  } else if (analysis_type == 7) {
    z_stim = c(0, 0)
  } else if (analysis_type == 8) {
    z_stim = c(0, 0)    
  }

# ------ Build plot
ggplot(data_file,
       aes(y = .data[[dv]],
           x = condition,
           col = session)) +
  geom_point(aes(y = .data[[dv]]),
             alpha = 0.5,
             size = 2,
             position = position_jitterdodge(jitter.width = 0.15,
                                             jitter.height = 0,
                                             dodge.width = 0.3)) +
  geom_boxplot(width = 0.3,
               alpha = 0,
               lwd = 1) +
  geom_flat_violin(aes(fill = session),
                   position = position_nudge(x = .2, y = 0),
                   alpha = .5) +
  
  labs(y = "z scores",
       x = "Movement Condition",
       color = "Session",
       fill = "Session",
       subtitle = "Behavioural data") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray") +
  annotate("segment", x = 0.6, xend = 1.4, y = z_stim[1], yend = z_stim[1], color = "darkgray", alpha = 0.5) +
  annotate("segment", x = 1.6, xend = 2.4, y = z_stim[2], yend = z_stim[2], color = "darkgray", alpha = 0.5) +
  scale_color_manual(values = met.brewer("Java", 2)) +
  scale_fill_manual(values = met.brewer("Java", 2)) +
  scale_x_discrete(labels = c("Three beat", "Four beat")) +
  apa7

# Save
name_file <- c("african_clap_zscores.png")
ggsave(paste(output_dir, name_file, sep = '/'), plot = last_plot(),
       dpi = 300, width = 35, height = 25, units = "cm", scale = 0.5)
```

# --------------------------------------
# Head movement

## Load data

```{r}
# ------ H1 ####
# ------ DATA LOADING
# ------ Indicate data path
path_z_snr = gsub("1_code/", "",
       here("2_output/data/4_final/head/z_snr/"))

# ------ List all files
list_files <- matrix(list.files(path = path_z_snr),
                    byrow = TRUE) 

# ------ Load data
data_file <- data.frame()

for (participant in 1:nrow(list_files)) {
   
  participant_file <-  read_csv(paste(path_z_snr, list_files[participant],
                                      sep = "/"),
                                show_col_types = FALSE)
   
   data_file <- rbind(data_file, participant_file)
}

# ------ Keep only relevant participants
data_file <- 
  data_file[which(data_file$participant %in% 
                    c(participant_pool$african_duple,
                      participant_pool$african_triple)), ]

# ------ Prepare data
# Change data format
data_file[1:4] <- lapply(data_file[1:4], as.factor)
```

## Stats

### Descriptives

```{r}
# ------ STATISTICS
# ------ Descriptive
stat_descr <-
  data_file %>%
  group_by(condition, session) %>%
  get_summary_stats(z_snr,
                    type = "mean_sd")
print(stat_descr)

```

### Assumptions checks

Normality of residuals assumption not met; but homoscedasticity assumption is met.
One outlier is detected. 

```{r}
# ------ Assumptions check
model <- lmer(as.formula("z_snr ~ condition * session + (1 | participant)"),
              data = data_file)

check_normality(model)
check_heteroscedasticity(model)

# ------ Identify outliers
check_outliers(model, method = "zscore")
```

#### Outliers correction

Outliers are corrected using Winsorisation (see [Leys et al., 2019](https://biblio.ugent.be/publication/8619726)).

```{r}
# ------ Deal with outliers
backup = data_file

data_file$z_snr <-
  Winsorize(data_file$z_snr,
            val = quantile(data_file$z_snr, probs = c(0, 0.99)))

check_outliers(model, method = "zscore")
```

#### Data transformation

```{r}
# ------ Transform data
model <- lmer(as.formula("z_snr ~ condition * session + (1 | participant)"),
              data = data_file)

# Square-root
data_file$sqrt_z_snr <- sqrt(data_file$z_snr + abs(min(data_file$z_snr)) + 1)
model_sqrt <- lmer(sqrt_z_snr ~ condition * session + (1 | participant),
                  data = data_file)
check_normality(model_sqrt)

```


### Inferential

No significant effect.

Note that in the GLM, dof = 76 (insted of 38) in previous models because the participant-level variance is zero (to test: > summary(model_sqrt)$varcor), meaning that including participant as a random effect doesn't explain any variation in z_snr.
This suggests that z_snr does not meaningfully vary across participants, making the random intercept redundant.
Since the random effect is zero, model_sqrt behaves as a fixed-effects model, increasing the denominator dof (DenDF) from 38 to 76.

```{r}
# ------ Inferential
# Omnibus test
data_file %>%
  anova_test(dv = sqrt_z_snr, wid = participant,
             between = condition,
             within = session,
             effect.size = "pes") %>%
  get_anova_table()

# GLM
anova(model_sqrt, type = 3) 
effectsize::eta_squared(model_sqrt)
```

## Plot

```{r}
# ------ PLOT
ggplot(data_file,
       aes(y = sqrt_z_snr,
           x = condition,
           col = session)) +
  geom_point(aes(y = sqrt_z_snr),
             alpha = 0.5,
             size = 2,
             position = position_jitterdodge(jitter.width = 0.15,
                                             jitter.height = 0,
                                             dodge.width = 0.3)) +
  geom_boxplot(width = 0.3,
               alpha = 0,
               lwd = 1) +
  geom_flat_violin(aes(fill = session),
                   position = position_nudge(x = .2, y = 0),
                   alpha = .5) +
  labs(y = "z scores",
       x = "Condition",
       color = "Session",
       fill = "Session",
       subtitle = "EEG data") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray") +
  geom_hline(yintercept = -0.329, color = "darkgray", alpha = 0.5) +
  scale_x_discrete(labels = c("Three-beat", "Four-beat")) +
  scale_y_continuous(name = "z scores", breaks = seq(from = -2, to = 5, by = 1), limits = c(-2, 5)) +
  apa7
```

# --------------------------------------
# Clapping (learning)

## Load data

```{r}
# ------ DATA LOADING
# ------ Indicate data path
path_asynch = gsub("1_code/", "",
       here("2_output/data/4_final/clap/asynchrony/"))

# ------ List all files
list_files <- matrix(list.files(path = path_asynch, pattern = "asynch.csv"),
                    byrow = TRUE) 

# ------ Load data
data_file <- data.frame()

for (participant in 1:nrow(list_files)) {
  
  file_path <- paste(path_asynch, list_files[participant], sep = "/")
  
  participant_file <- read_delim(file_path,
                                 delim = guess_delim(file_path), 
                                 show_col_types = FALSE)
  
  participant_row <- participant_file[1, 1:3]
  
  participant_row$signed_asynchrony <- 
    mean(participant_file$signed_asynchrony, na.rm = TRUE)
  
  participant_row$r_asynch <- 
    mean(participant_file$r_asynch, na.rm = TRUE)  
  
  data_file <- rbind(data_file, participant_row)
  
}

# ------ Keep only relevant participants
data_file <- 
  data_file[which(data_file$participant %in% 
                    c(participant_pool$african_duple,
                      participant_pool$african_triple)), ]

# ------ Prepare data
# Change data format
data_file[1:3] <- lapply(data_file[1:3], as.factor)
```


## Stats

### Descriptives

```{r}
# ------ STATISTICS
# ------ Descriptive
stat_descr <-
  data_file %>%
  group_by(condition) %>%
  get_summary_stats(signed_asynchrony,
                    type = "mean_sd")
print(stat_descr)

```

### Assumptions checks

Normality of residuals and homoscedasticity assumptions are met.
No outlier is detected.

```{r}
# ------ Assumptions check
model <- lm(signed_asynchrony ~ condition,
              data = data_file)

check_normality(model)
check_heteroscedasticity(model)

# ------ Identify outliers
check_outliers(model, method = "zscore")

```

### Inferential
No significant effect.

```{r}
# ------ Inferential
# Omnibus test
data_file %>%
  anova_test(dv = signed_asynchrony, wid = participant,
             between = condition,
             effect.size = "pes") %>%
  get_anova_table()

# TOST
data_wide <- data_file %>%
  pivot_wider(
    names_from = condition,
    values_from = signed_asynchrony,
    names_prefix = "condition_"
  )

tsum_TOST(m1 = mean(data_wide$condition_2, na.rm = TRUE),
          sd1 = sd(data_wide$condition_2, na.rm = TRUE),
          n1 = 20,
          m2 = mean(data_wide$condition_3, na.rm = TRUE),
          sd2 = sd(data_wide$condition_3, na.rm = TRUE),
          n2 = 20,
          r12 = cor(data_wide$condition_2[1:20], data_wide$condition_3[21:40]),
          hypothesis = "EQU",
          paired = FALSE,
          eqb = 0.01, # based on Rose et al. (2021)
          eqbound_type = "raw")

# GLM
anova(model)
effectsize::eta_squared(model)
```

## Plot

```{r}
# ------ PLOT
ggplot(data_file,
       aes(y = signed_asynchrony,
           x = condition,
           col = condition)) +
  geom_point(aes(y = signed_asynchrony),
             alpha = 0.5,
             size = 2,
             position = position_jitterdodge(jitter.width = 0.15,
                                             jitter.height = 0,
                                             dodge.width = 0.3)) +
  geom_boxplot(width = 0.3,
               alpha = 0,
               lwd = 1) +
  geom_flat_violin(aes(fill = condition),
                   position = position_nudge(x = .2, y = 0),
                   alpha = .5) +
  labs(y = "Signed Asynchrony",
       x = "Condition") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray") +
  scale_x_discrete(labels = c("Three-beat", "Four-beat")) +
  guides(color = "none", fill = "none") +
  apa7
```

# Clapping (learning stability)
## Stats

### Descriptives

```{r}
# ------ STATISTICS
# ------ Descriptive
stat_descr <-
  data_file %>%
  group_by(condition) %>%
  get_summary_stats(r_asynch,
                    type = "mean_sd")
print(stat_descr)

```
### Assumptions checks

Normality of residuals is not met. 
Homoscedasticity assumption is met.
One outlier is detected.

```{r}
# ------ Assumptions check
model <- lm(r_asynch ~ condition,
              data = data_file)

check_normality(model)
check_heteroscedasticity(model)

# ------ Identify outliers
check_outliers(model, method = "zscore")

```

#### Outliers correction

Outlier is corrected using Winsorisation (see [Leys et al., 2019](https://biblio.ugent.be/publication/8619726)).

```{r}
# ------ Deal with outliers
data_file$r_asynch <-
  Winsorize(data_file$r_asynch,
            val = quantile(data_file$r_asynch, probs = c(0.02, 1)))

model <- lm(r_asynch ~ condition, data = data_file)
check_outliers(model, method = "zscore")
check_normality(model)
check_heteroscedasticity(model)
```

#### Data transformation
The data is transformed using a rank-based inverse normal transformation. 
```{r}
# ------ Transform data

# Rank-based inverse normal transformation 
data_file <- data_file %>%
  mutate(r_asynch_norm = orderNorm(r_asynch)$x.t)

model_rint <- lm(r_asynch_norm ~ condition, data = data_file)
check_normality(model_rint)

```

### Inferential

No significant effect.

```{r}
# ------ Inferential
# Omnibus test
data_file %>%
  anova_test(dv = r_asynch_norm, wid = participant,
             between = condition,
             effect.size = "pes") %>%
  get_anova_table()

```

## Plot

```{r}
# ------ PLOT
ggplot(data_file,
       aes(y = r_asynch,
           x = condition,
           col = condition)) +
  geom_point(aes(y = r_asynch),
             alpha = 0.5,
             size = 2,
             position = position_jitterdodge(jitter.width = 0.15,
                                             jitter.height = 0,
                                             dodge.width = 0.3)) +
  geom_boxplot(width = 0.3,
               alpha = 0,
               lwd = 1) +
  geom_flat_violin(aes(fill = condition),
                   position = position_nudge(x = .2, y = 0),
                   alpha = .5) +
  labs(y = "Vector Strength of the Asynchrony",
       x = "Condition") +
  #geom_hline(yintercept = 0, linetype = "dashed", color = "gray") +
  scale_x_discrete(labels = c("Three-beat", "Four-beat")) +
  guides(color = "none", fill = "none") +
  apa7
```


# --------------------------------------
# Stepping (learning)

## Load data

```{r}
# ------ DATA LOADING
# ------ Indicate data path
path_asynch = gsub("1_code/", "",
       here("2_output/data/4_final/step/asynchrony/"))

# ------ List all files
list_files <- matrix(list.files(path = path_asynch, pattern = ".csv"),
                    byrow = TRUE) 

# ------ Load data
data_file <- data.frame()

for (participant in 1:nrow(list_files)) {
  
  file_path <- paste(path_asynch, list_files[participant], sep = "/")
  
  participant_file <- read_delim(file_path,
                                 delim = guess_delim(file_path), 
                                 show_col_types = FALSE)
  
  participant_row <- participant_file[1, 1:3]
  
  participant_row$signed_asynchrony <- 
    mean(participant_file$signed_asynchrony, na.rm = TRUE)
  
  participant_row$r_asynch <- 
    mean(participant_file$r_asynch, na.rm = TRUE)  
  
  data_file <- rbind(data_file, participant_row)
  
}

# ------ Keep only relevant participants
data_file <- 
  data_file[which(data_file$participant %in% 
                    c(participant_pool$african_duple,
                      participant_pool$african_triple)), ]

# ------ Prepare data
# Change data format
data_file[1:3] <- lapply(data_file[1:3], as.factor)
```


## Stats

### Descriptives

```{r}
# ------ STATISTICS
# ------ Descriptive
stat_descr <-
  data_file %>%
  group_by(condition) %>%
  get_summary_stats(signed_asynchrony,
                    type = "mean_sd")
print(stat_descr)

```

### Assumptions checks

Normality of residuals and homoscedasticity assumptions are met.
No outlier is detected.

```{r}
# ------ Assumptions check
model <- lm(signed_asynchrony ~ condition,
              data = data_file)

check_normality(model)
check_heteroscedasticity(model)

# ------ Identify outliers
check_outliers(model, method = "zscore")
```

### Inferential

No significant effect.

```{r}
# ------ Inferential
# Omnibus test
data_file %>%
  anova_test(dv = signed_asynchrony, wid = participant,
             between = condition,
             effect.size = "pes") %>%
  get_anova_table()

# TOST
data_wide <- data_file %>%
  pivot_wider(
    names_from = condition,
    values_from = signed_asynchrony,
    names_prefix = "condition_"
  )

tsum_TOST(m1 = mean(data_wide$condition_2, na.rm = TRUE),
          sd1 = sd(data_wide$condition_2, na.rm = TRUE),
          n1 = 20,
          m2 = mean(data_wide$condition_3, na.rm = TRUE),
          sd2 = sd(data_wide$condition_3, na.rm = TRUE),
          n2 = 20,
          r12 = cor(data_wide$condition_2[1:20], data_wide$condition_3[21:40]),
          hypothesis = "EQU",
          paired = FALSE,
          eqb = 0.01, # based on Rose et al. (2021)
          eqbound_type = "raw")

# GLM
anova(model)
effectsize::eta_squared(model)
```

## Plot

```{r}
# ------ PLOT
ggplot(data_file,
       aes(y = signed_asynchrony,
           x = condition,
           col = condition)) +
  geom_point(aes(y = signed_asynchrony),
             alpha = 0.5,
             size = 2,
             position = position_jitterdodge(jitter.width = 0.15,
                                             jitter.height = 0,
                                             dodge.width = 0.3)) +
  geom_boxplot(width = 0.3,
               alpha = 0,
               lwd = 1) +
  geom_flat_violin(aes(fill = condition),
                   position = position_nudge(x = .2, y = 0),
                   alpha = .5) +
  labs(y = "Signed Asynchrony",
       x = "Condition") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray") +
  scale_x_discrete(labels = c("Three-beat", "Four-beat")) +
  guides(color = "none", fill = "none") +
  apa7
```

# Stepping (learning stability)
## Stats

### Descriptives

```{r}
# ------ STATISTICS
# ------ Descriptive
stat_descr <-
  data_file %>%
  group_by(condition) %>%
  get_summary_stats(r_asynch,
                    type = "mean_sd")
print(stat_descr)

```

### Assumptions checks

Normality of residuals and homoscedasticity assumptions are not met. 
One outlier is detected.

```{r}
# ------ Assumptions check
model <- lm(r_asynch ~ condition,
              data = data_file)

check_normality(model)
check_heteroscedasticity(model)

# ------ Identify outliers
check_outliers(model, method = "zscore")

```

#### Outliers correction

Outlier is corrected using Winsorisation (see [Leys et al., 2019](https://biblio.ugent.be/publication/8619726)).

```{r}
# ------ Deal with outliers
data_file$r_asynch <-
  Winsorize(data_file$r_asynch,
            val = quantile(data_file$r_asynch, probs = c(0.03, 1), na.rm = TRUE))

model <- lm(r_asynch ~ condition, data = data_file)
check_outliers(model, method = "zscore")
check_normality(model)
check_heteroscedasticity(model)
```

#### Data transformation
The data is transformed using a rank-based inverse normal transformation. 
```{r}
# ------ Transform data

# Rank-based inverse normal transformation 
data_file <- data_file %>%
  mutate(r_asynch_norm = orderNorm(r_asynch)$x.t)

model_rint <- lm(r_asynch_norm ~ condition, data = data_file)
check_normality(model_rint)

```

### Inferential

No significant effect.

```{r}
# ------ Inferential
# Omnibus test
data_file %>%
  anova_test(dv = r_asynch_norm, wid = participant,
             between = condition,
             effect.size = "pes") %>%
  get_anova_table()

```

## Plot

```{r}
# ------ PLOT
ggplot(data_file,
       aes(y = r_asynch_norm,
           x = condition,
           col = condition)) +
  geom_point(aes(y = r_asynch_norm),
             alpha = 0.5,
             size = 2,
             position = position_jitterdodge(jitter.width = 0.15,
                                             jitter.height = 0,
                                             dodge.width = 0.3)) +
  geom_boxplot(width = 0.3,
               alpha = 0,
               lwd = 1) +
  geom_flat_violin(aes(fill = condition),
                   position = position_nudge(x = .2, y = 0),
                   alpha = .5) +
  labs(y = "Vector Strength of the Asynchrony",
       x = "Condition") +
  #geom_hline(yintercept = 0, linetype = "dashed", color = "gray") +
  scale_x_discrete(labels = c("Three-beat", "Four-beat")) +
  guides(color = "none", fill = "none") +
  apa7
```



